---
title: "Fear and Foraging"
author: "Kasia Wojczulanis-Jakubas"
date: "8 wrzeœnia 2017"
output: html_document
---
```{r, echo = FALSE, message = FALSE, warning = FALSE, results='hide'}

library(dplyr)
library(ggplot2)

# library(FactoMineR)
# library(factoextra)

library (lme4) 
# library (lmerTest) 

# Download data, filtering (to kepp only contr which for there is respective experiment)
# and set working directory for the output

# Data set for latency
setwd("D:\\JA\\Colibri\\Ms_Fear and Foraging\\Data\\Personality exp_latency")
lacy <- read.csv("latency.csv", header = TRUE, sep = ",")
lacy <- lacy %>% 
  dplyr::filter(Conf != "Ctrl0") %>%
  dplyr::filter(Conf != "Ctrl3") %>%
  gdata::drop.levels()

# Data set for exploration, risk avoidance and risk arousal
setwd("D:\\JA\\Colibri\\Ms_Fear and Foraging\\Data\\Personality exp_explor_risk_arousal")
persimpl <- read.csv("persIDc.csv", header = TRUE, sep = ",")
persimpl <- persimpl %>% 
  dplyr::filter(Configuration != "Ctrl0") %>%
  dplyr::filter(Configuration != "Ctrl3") %>%
  gdata::drop.levels()

# Working directory for the analysis output 
setwd("D:\\JA\\Colibri\\Ms_Fear and Foraging\\Analysis")


```

## Risk avoidance 1

Latency in approach to the feeder (i.e. the time between appearence in front of the feeder and onset of foraging. Modelled in regard to the conditions (fixed factor; control vs experimental), with birds ID as a random factor

```{r, echo = FALSE, message=FALSE, warning=FALSE}

########### Modelling  ############
# names(lacy) # dataset
# modelRA1 <- lmer(Latency ~ ConfSimpl + (1|Colcode), data = lacy)
# plot(modelRA1)
modelRA1 <- lmer(log(Latency) ~ ConfSimpl + (1|Colcode), data = lacy)
summary(modelRA1)
estimRA1 <- modelRA1@beta[2]
lowcfRA1 <- confint(modelRA1, level = .90)[4,1]
uppcfRA1 <- confint(modelRA1, level = .90)[4,2]

########### Plot  #################
ggplot(lacy, aes(y = Latency, x = ConfSimpl)) +
geom_boxplot() +
    theme_bw() +
    scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
    scale_y_continuous(name = "Latency [secs]")

    
```


## Risk avoidance 2  
Two "competing" variables: stanouts (number of breaks during the foraging, standardized by the total time at the feeder) and mean_durouts (mean duration of breaks; mean calculated per foraging visit of an idividual).

``` {r, echo = FALSE, message = FALSE, warning = FALSE}

########### Modelling  ############
# names(persimpl) # dataset
modelRA2a <- lmer(stan_nouts ~ treat + (1|Color.code), data = persimpl)
# plot(modelRA2a)
summary(modelRA2a)
estimRA2a <- modelRA2a@beta[2]
lowcfRA2a <- confint(modelRA2a, level = .90)[4,1]
uppcfRA2a <- confint(modelRA2a, level = .90)[4,2]

modelRA2b <- lmer(mean_durouts ~ treat + (1|Color.code), data = persimpl)
# plot(modelRA2b)
summary(modelRA2b)
estimRA2b <- modelRA2b@beta[2]
lowcfRA2b <- confint(modelRA2b, level = .90)[4,1]
uppcfRA2b <- confint(modelRA2b, level = .90)[4,2]

########### Plot  ############
ggplot(persimpl, aes(y = mean_durouts, x = treat)) +
  geom_boxplot() +
  theme_bw() +
  scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
  scale_y_continuous(name = "Standardized duration of breaks during the foraging")
 


ggplot(persimpl, aes(y = stan_nouts, x = treat)) +
  geom_boxplot() +
  theme_bw() +
  scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
  scale_y_continuous(name = "Standardized number of breaks during the foraging")
 

```


## Exploration   
Two variables "competitive" variables:stannchange (number of changes of the feeder) and stannflo (number of diferent feeders used), both standardized by the total time spent at the feeder.


``` {r, echo = FALSE, message = FALSE, warning = FALSE}

########### Modelling  ############
# names(persimpl) # dataset
# modelEa <- lmer(stan_nflo ~ treat + (1|Color.code), data = persimpl)
# plot(modelEa)
modelEa <- lmer(log(stan_nflo+1) ~ treat + (1|Color.code), data = persimpl)
# plot(modelEa)
summary(modelEa)
estimEa <- modelEa@beta[2]
lowcfEa <- confint(modelEa, level = .90)[4,1]
uppcfEa <- confint(modelEa, level = .90)[4,2]

modelEb <- lmer(log(stan_nflochang+1) ~ treat + (1|Color.code), data = persimpl)
# plot(modelEb)
summary(modelEb)
estimEb <- modelEb@beta[2]
lowcfEb <- confint(modelEb, level = .90)[4,1]
uppcfEb <- confint(modelEb, level = .90)[4,2]

########### Plot  ############
ggplot(persimpl, aes(y = stan_nflo, x = treat)) +
# ggplot(persimpl, aes(y = stan_nflochang, x = treat)) +
    geom_boxplot() +
    theme_bw() +
    scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
    scale_y_continuous(name = "Standardized number of feeders used")
    # scale_y_continuous(name = "Standardized number of feeders change")
 


ggplot(persimpl, aes(y = stan_nflochang, x = treat)) +
# ggplot(persimpl, aes(y = stan_nflochang, x = treat)) +
    geom_boxplot() +
    theme_bw() +
    scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
    scale_y_continuous(name = "Standardized number of feeder changes")
    # scale_y_continuous(name = "Standardized number of feeders change")
 

```


## Total energy intake
Total duation of feeding (a proxy for energy intake) – analysed in regard to the configuration, with birds ID included in the model a a random effect.

```{r, echo = FALSE, warning=FALSE, message = FALSE}
  
########### Modelling  ############
# names(persimpl) # dataset
modelTE <- lmer(log(tot_durins) ~ treat + (1|Color.code), data = persimpl)
# plot(modelTE)
summary(modelTE)  
estimTE <- modelTE@beta[2]
lowcfTE <- confint(modelTE, level = .90)[4,1]
uppcfTE <- confint(modelTE, level = .90)[4,2]

########### Plot  ############
ggplot(persimpl, aes(x = treat, y= tot_durins)) + geom_boxplot() +
    theme_bw() +
    scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
    scale_y_continuous(name = "Total duration of the feeding [secs]")

```

## Arousal

Three traits: ferocity, speed and total distance. PCA performed to see if the three could be connected. 

```{r, echo = FALSE, warning=FALSE, message = FALSE}
  
########### PCA  ############
# names(persimpl) # dataset
# perspca <- persimpl [,12:14] # df need to consist of quatitive variables only
# res.pca <- PCA(perspca, graph = T)
# res.pca$eig # proportions of the retained varaitions 

########### Modelling  ############
# names(persimpl) # dataset
modelA <- lmer(log(mov_feroc) ~ treat + (1|Color.code), data = persimpl)
# plot(modelA)
summary(modelA)  
estimA <- modelA@beta[2]
lowcfA <- confint(modelA, level = .90)[4,1]
uppcfA <- confint(modelA, level = .90)[4,2]

library(dplyr)
sumtab <- persimpl %>% group_by(treat) %>% 
  summarize(mean_mov_feroc = mean(mov_feroc))

  sumtab$mean_mov_feroc[2] - sumtab$mean_mov_feroc[1]

########### Plot  ############
ggplot(persimpl, aes(x = treat, y= mov_feroc)) + geom_boxplot() +
    theme_bw() +
    scale_x_discrete(labels = c("Control",  "Experimental"), name = "Conditions")+
    scale_y_continuous(name = "Ferocity")

```


## Model esitmates for all variables

```{r, echo = FALSE, message=FALSE, warning=FALSE}

label <- c("risk-av-durlatency", "risk-av-nbreaks", "risk-av-durbreaks", "explor-nfchang", "feroc", "energyintake")
estim <- c(estimRA1, estimRA2a, estimRA2b, estimEb, estimA, estimTE)
uppcf  <- c(uppcfRA1, uppcfRA2a, uppcfRA2b, uppcfEb, uppcfA, uppcfTE) 
lowcf  <- c(lowcfRA1, lowcfRA2a, lowcfRA2b, lowcfEb, lowcfA, lowcfTE) 

df <- data.frame(label, estim, lowcf, uppcf)
df$label <- factor(df$label, levels=rev(df$label))

# Plot
ggplot(data=df, aes(x=label, y=estim, ymin=lowcf, ymax=uppcf)) +
        geom_pointrange() + 
        coord_flip() +
        geom_hline(yintercept=0, lty=2) +  
        xlab("Label") + ylab("Model estimates and 90% CI") +
        theme_bw()  


```


