---
title: "Hermit's personality - data preparation"
author: "Kasia Wojczulanis-Jakubas"
date: "16/05/2016"
output: html_document
---
##----------------------------------------------------------------------------------------------------------------------
#### Some auxillary functions

*To create a temporary variable for subsequent calculations/subsetting*


``` {r, echo = FALSE, message  = FALSE, warning=FALSE}


shcut1    <- "numfd"
fullname1 <- "numbers of feeders" 
def1      <-  "recodes feeders and intervals (A, B, C, OUT -> 1, 2, 3, 0)"   

shcut2    <- "seqfd"
fullname2 <- "sequential numbers of feeders" 
def2      <- "sequentially numbers feeders and intervals" 

shcut3    <- "selou"
fullname3 <- "selection of external OUTS"
def3      <- "indicates records to delete (external OUTS)"

shcut4    <- "subsdf"
fullname4 <- "subsetting data frames"
def4      <- "to subet valid records (no external OUTS)"

Short_cut <- c(shcut1, shcut2, shcut3, shcut4)
Full_name <- c(fullname1, fullname2,fullname3,fullname4)
Definition <- c(def1, def2, def3, def4)

auxfun <- as.data.frame(cbind(Short_cut, Full_name, Definition))

library(ascii)
tab1 <- ascii(auxfun)
print (tab1)

```

```{r}

numfd <- function (X1) {
  for (i in 1:length(X1$Feeder)) {
    if (X1$Feeder[i] == "A") {
      X1$feeder[i] = 1
    } else {
      if (X1$Feeder[i] == "B") {
        X1$feeder[i] = 2 
      } else {
        if (X1$Feeder[i] == "C") {
          X1$feeder[i] = 3 
        } else {
          X1$feeder[i] = 0
        }
      }
    }
  }
  result <- X1 
  return(result)
} 

```

```{r}

seqfd <- function (Y1) {
  Y1$seqFeeder <- rep(x = seq_along(rle(Y1$feeder)$lengths), times = rle(Y1$feeder)$lengths)
  return(Y1)
} 

```

```{r}

selout <- function (Z1) {
  lastrow <-length(rle(Z1$feeder)$lengths) 
  Z1$sel <-  ifelse (Z1$seqFeeder == lastrow & Z1$Feeder == "OUT" | Z1$seqFeeder == 1 & Z1$Feeder == "OUT",
                           "to delete", "OK")
  return(Z1)
} 

```

```{r}
subsdf<- function (S1) {
  S1 <-S1[S1$sel == "OK", ]
}

```


#### Loading and processing all data frames with auxillary functions


```{r}

library(plyr)
setwd("D:/JA/Colibri/Personality tests 2015/Pers.Video cuts 2015/csv tracker files")


dataframes <- list.files(path=getwd())  
list_dataframes <- llply(dataframes, read.csv, header = T, sep = ";", dec = ",") # list of all data frames

list_dataframes <- llply(list_dataframes, numfd)
list_dataframes <- llply(list_dataframes, seqfd)
list_dataframes <- llply(list_dataframes, selout)
list_dataframes <- llply(list_dataframes, subsdf)

```


#------------------------------------------------------------
# Personality variables


**Number of feeders used**

*Absolute number of feeders = number of different letters; e.g A:B, B:A, A:B = 2; OUT's need to be excluded.*

*A proxy for explorative trait.*
*The HIGER value, the MORE explorative individual.*


```{r}

nflo <- function (C1) {
  if (any(C1$Feeder == "OUT")){
    abs_nflo<- length(unique(C1$Feeder)) - 1
  } else {abs_nflo<- length(unique(C1$Feeder))}
}

pers1 <- ldply(list_dataframes, nflo)

```



**Number of feeders' changes**

*eg. A:B, B:A, A:B =  3; internal OUTS are redundant at this stage - requires a temporary dataset.*

*A proxy for explorative trait, may be stronger/better than abolute number of feeders.*
*The HIGHER value, the MORE explorative individual.*


```{r}

subsdf2<- function (F1) { # function to subset
  F1 <-F1[F1$Feeder != "OUT", ] 
}


list_dataframes1 <- llply(list_dataframes, subsdf2) # subsetting


nchang <- function (H1){ # function to calculate the number of changes
  nflo_chang <- length(rle(as.numeric(H1$Feeder))$values)-1 
}


pers2 <- ldply(list_dataframes1, nchang)

```



**Number of intra-feeding intervals**

*Number of ‘OUTS’ in between; e.g. A-OUT-B-OUT-C = 2.*

*A proxy for risk-avoidance trait, partly also for general arousal.*
*The HIGHER value, the MORE risk-avoiding individual, with higher arousal.*


```{r}

nout<- function (G1) {
  nouts<-length(which(rle(G1$feeder)$values == 0))
  return(nouts)
}

pers3 <- ldply(list_dataframes, nout)

```



**Number of feeding intervals**

*Number of ‘INS’ (all feeders, i.e. A, B, B).*

*A proxy for risk-avoidance trait, partly also for general arousal; kind of "reverse"" of the number of inter-feeding intervals.* 
*The HIGHER value, the MORE risk-avoiding individual, with higher arousal.*

```{r}

pers4 <- as.data.frame(pers3$V1+1)

```



**Feeding duration - mean and total feeding duration (in seconds)**

*Mean: a simple mean duration of 'INS' intervals. Total: sum of all ‘INS’ durations;  e.g.  A+B+C.* 
*Single frame duration - 0.33 sec; requires temp data set: list_dataframes1 - no internal 'OUTS'.*

*A proxy for risk-avoidance trait, partly also for general arousal; kind of "reverse"" of the number of inter-feeding intervals.* 
*the HIGHER value, the LESS risk-avoiding individual, with LOWER arousal.*


```{r}

require(stats)

dins <- function (T1){
  tapply(T1$Feeder, T1$seqFeeder, length) # need to be run on df without internal 'OUTS'
}

tempdf1 <- llply(list_dataframes1, dins)  
pers5 <- ldply(tempdf1, mean)
pers5$V1 <-  pers5$V1*0.033


pers6 <- ldply(tempdf1, sum)
pers6$V1 <-  pers6$V1*0.033


```



**Duration of interfeeding intervals  - mean and total  (in seconds)**

*Mean: a simple mean duration of 'OUTS' intervals. Total: sum of all ‘OUTS' in between;  e.g.  A-from OUT1-B-OUT 2-C ->  OUT1+OUT2.*
*Single frame duration - 0.33 sec; requires temp data set - only 'OUTS' records.*

*A proxy for risk-avoidance trait, partly also for general arousal; kind of "reverse"" of the number of inter-feeding intervals.* 
*The HIGHER value, the MORE risk-avoiding individual, with HIGHER arousal.*


```{r}

subsdf3<- function (J1) { # function to subset
  J1 <-J1[J1$Feeder == "OUT", ] 
}

list_dataframes2 <- llply(list_dataframes, subsdf3) # subsetting

require(stats)

douts <- function (U1){
  tapply(U1$Feeder, U1$seqFeeder, length)
}

tempdf2 <- llply(list_dataframes2, douts)
pers7 <- ldply(tempdf2, mean) # applying the function to the proper data set, calcualting the mean

pers7[is.na(pers7)] <- 0 # NA values need to be replaced with 0
pers7$V1 <-  pers7$V1*0.033 

pers7a <- ldply(tempdf2, sum) # calculating the total
pers7a$V1 <- pers7a$V1*0.033

```



**Calculation of duration of foraging**

*Sum of duration of all 'INS' and 'OUTS'.*
* More like an outcome of the personality, to be used in further analyses.*


```{r}

pers8 <- as.data.frame(pers6$V1+ pers7a$V1)

```



**Space usage - total distance, 'speed' and ferocity**

*A proxies for general arousal; Three components - a) total distance covered during the foraging, b) 'speed' (rather distance standardized by total foraging time), and c) ferocity (coeficient variance of deviations from the feeders)*
*The HIGHER values, the  HIGHER general arousal.*


```{r}

spcals<-function(X){ # total distance
  di<-NULL
  for(i in 2:nrow(X))
  {lx<-X$x[i]-X$x[i-1]
  ly<-X$y[i]-X$y[i-1]  
  di[i-1]<-sqrt(lx^2+ly^2)
  }
  return(sum(di))
}


spcalmean<-function(X){ # mean spead (frame = row = 0.033 sec)
  di<-NULL
  for(i in 2:nrow(X))
  {lx<-X$x[i]-X$x[i-1]
  ly<-X$y[i]-X$y[i-1]  
  di[i-1]<-sqrt(lx^2+ly^2)
  }
  return(mean(di)/(nrow(X)*0.033))
}

# An auxillary function to calculate coeficent of variance
CV <- function(z){
  sd(z)/mean(z)
}

spcalcv<-function(X){ # express ferocity
  di<-NULL
  for(i in 2:nrow(X))
  {lx<-X$x[i]-X$x[i-1]
  ly<-X$y[i]-X$y[i-1]  
  di[i-1]<-sqrt(lx^2+ly^2)
  }
  return(CV(di))
}


pers9a <- ldply(list_dataframes, spcals)
pers9b <- ldply(list_dataframes, spcalmean)
pers9c <- ldply(list_dataframes, spcalcv)

```


#------------------------------------------------------------
## Personality variables - data frame

*Putting all variables into a single data frame, to be merged with birds IDs*


```{r}

pers <-cbind(pers1, pers2, pers3, pers4, pers5, pers6, pers7, pers7a, pers8, pers9a, pers9b, pers9c)
colnames(pers) <- c("abs_nflo", "nflo_chang", "nouts", "nins", "mean_durins", "tot_durins", "mean_durouts", 
                    "tot_durouts","tot_durfor", "mov_totdist",  "mov_spead", "mov_feroc") 


pers$Video.cut <- dataframes
pers$treat<-substr(pers$Video.cut, 38, 40)

```

#------------------------------------------------------------
#### Birds ID 

```{r}
setwd("D:/JA/Colibri/video cuts 2015/Pers.Video cuts 2015") 
birdsID <- read.csv (file = "detailed data.csv", header = T, sep = ";") # detailed data - df with the birds Id
```
  
  
``` {r, echo = FALSE, eval = FALSE}

# Exploring data set
names(birdsID)
str(birdsID)
sum(is.na(birdsID$ID..)) # to check if all birds ID are provided
sum(is.na(birdsID$Video.cut)) # to check if all video ID are provided

# To merge the two data frame
# Adjusting Video cuts names

# Check for n characters
# 
# length(pers$Video.cut)
# class(pers$Video.cut)

# pers$Video.cut # for pers data frame
# for (p in 1:119) {
#   nID <- nchar(pers$Video.cut)
# }
# nID
# 
# 
# length(birdsID$Video.cut)
# class(birdsID$Video.cut)
# 
# for (f in 1:151) {  # for birds Id data frame
#   nId <- nchar(as.character(birdsID$Video.cut))
# }
# nId

```

#### Merging personality variables data frame (pers) with birds ID

```{r}

Idtemp1 <- as.character(birdsID$Video.cut) #Adjusting Video.cut in birdsID data frame
Idtemp2 <- paste(Idtemp1, ".csv", sep = "")
birdsID$Video.cut <-Idtemp2         


# merging the two data frames
persID <- merge(pers, birdsID, by.x = "Video.cut", all.x = TRUE)

```


``` {r, echo = FALSE, eval = FALSE}

# Inspection of the merged data frames
View(persID)
head(persID)
length(persID$Video.cut)
names(persID)
sum(!complete.cases(persID))  # counts number of missing values
which(!complete.cases(persID))  # counts number of missing values

```

#### Cleanig up data 

```{r}

doubts <- which(!persID$Comments == "")
bedrecc <- doubts[c(1:3, 11,12)] # list of records that behaviour could be changed by an external factores (bee/bird)

persIDc <- persID[-bedrecc,-c(15:18, 20:22, 25:28)] # data frame without suspicious records, and redudunt columns
persIDc$ID.. <- as.factor(persIDc$ID..)
```


#### Personality variables transformation/standarization 

```{r}

persIDc$stan_nflo      <- persIDc$abs_nflo/persIDc$tot_durfor 
persIDc$stan_nflochang <- persIDc$nflo_chang/persIDc$tot_durfor
persIDc$stan_nouts     <- persIDc$nouts/persIDc$tot_durfor
persIDc$stan_nins      <- persIDc$nins/persIDc$tot_durfor
persIDc$stand_totdist  <- persIDc$mov_totdist/(persIDc$nflo_chang+1)/(persIDc$nouts+1) 

```



```{r}

write.csv(persIDc, file = "persIDc.csv") # personality data frame with birds ID, cleaned-up, ready-to-use 

```
