---
title: "Personality_data analysis"
author: "Kasia Wojczulanis-Jakubas"
date: "30/05/2016"
output: html_document
---

#------------------------------------------------------------
#### Data frames   
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(Rcpp)
library(gridExtra)
library(ggplot2)


```

```{r, echo = FALSE, message=FALSE, warning=FALSE}

setwd("D:/JA/Colibri/video cuts 2015/Pers.Video cuts 2015") 
persIDc <- read.csv (file = "persIDc.csv", header = T, sep = ",") # df with the pers vars, birds Id; cleaned up

persIDcs <- persIDc[, c(18, 15, 16, 19:22, 7:11, 13, 14, 23)] # df with selected columns
persIDcs$ID.. <- as.factor(persIDcs$ID..)


library(reshape)
melt_persIDcs <- melt(persIDcs, id=c("treat", "Configuration", "ID..")) 
melt_persIDcs_c <- melt_persIDcs[melt_persIDcs$treat == "Ctr" , ] 
melt_persIDcs_e <- melt_persIDcs[melt_persIDcs$treat == "Exp" , ] 


library(plyr)

persIDcs_m <- ddply(persIDcs, c("treat", "ID.."), summarise,
               mean_durins    = mean(mean_durins),
               tot_durins    = mean(tot_durins),
               mean_durouts    = mean(mean_durouts),
               tot_durouts    = mean(tot_durouts),
               tot_durfor    = mean(tot_durfor),
               mov_spead = mean(mov_spead),
               mov_feroc = mean(mov_feroc),
               stan_nflo = mean(stan_nflo),
               stan_nflochang = mean(stan_nflochang),
               stan_nouts = mean(stan_nouts),
               stan_nins = mean(stan_nins),
               stand_totdist = mean(stand_totdist)
)

melt_persIDcs_mc <- persIDcs_m[persIDcs_m$treat == "Ctr" , ] 
melt_persIDcs_me <- persIDcs_m[persIDcs_m$treat == "Exp" , ] 

```

persIDc # df with the pers vars, birds Id; cleaned up  
persIDcs # df with selected columns    
melt_persIDcs_c # melted df (by var), control    
melt_persIDcs_e # melted df (by var), experiment    
persIDcs_m # df with means ("treat", "ID..")     
melt_persIDcs_mc # melted means df (by var), control     
melt_persIDcs_me # melted means df (by var), experiment        


#------------------------------------------------------------

## Exploring pers variables   


#### Density plots for all pers variables - **CONTROL**    

```{r, fig.width=2.5, fig.height=2.5, message=FALSE, warning=FALSE}

library(ggplot2)


AllVard <- unique(melt_persIDcs_c$variable)
for (i in 1:length(AllVard))
{
  Var.i <- AllVard[i]
  dfs.i <- melt_persIDcs_c[melt_persIDcs_c$variable == Var.i, ]
  print(ggplot(data = dfs.i, aes(x = value)) + geom_density(aes(x = value), fill = "grey") + ggtitle(paste(Var.i, "_CTR",sep = "")) + theme(legend.position="none")) 
          
}

```  
  
  
#### Density plots for all pers variables - **EXPERIMENT**       
  
  
```{r, fig.width=2.5, fig.height=2.5, message=FALSE, warning=FALSE}

library(ggplot2)

AllVard1 <- unique(melt_persIDcs_e$variable)
for (i in 1:length(AllVard1))
{
  Var.i <- AllVard1[i]
  dfs.i <- melt_persIDcs_e[melt_persIDcs_e$variable == Var.i, ]
  print(ggplot(data = dfs.i, aes(x = value)) + geom_density(aes(x = value), fill = "grey") +
          ggtitle(paste(Var.i, "_EXP",sep = "")) + theme(legend.position="none")) 
          
  }

```



#### **Ctr vs exp** distributions of the pers variables - KS test      



```{r, results="hide", message=FALSE, warning=FALSE}

AllVard2c <- unique(melt_persIDcs_c$variable)
AllVard2e <- unique(melt_persIDcs_e$variable)

for (i in 1:length(AllVard2c))
{
  Varc.i <- AllVard2c[i]
  Vare.i <- AllVard2e[i]
  
  dfsc.i <- melt_persIDcs_c[melt_persIDcs_c$variable == Varc.i, ]
  dfse.i <- melt_persIDcs_e[melt_persIDcs_e$variable == Vare.i, ]
  print(Varc.i)
  print(ks.test(dfsc.i$value,   dfse.i$value, alternative = "two.sided"))  
  }

```

=> Concl ks.test: signif diff in mean_durouts, and mov_feroc    



#### Individual distribution by pers_var **CONTROL**         


 
``` {r, fig.width=2.5, fig.height=2.5, message=FALSE, warning=FALSE}

AllVar3 <- unique(melt_persIDcs_c$variable)
for (i in 1:length(AllVar3))
{
  Var.i <- AllVar3[i]
  dfs.i <- melt_persIDcs_c[melt_persIDcs_c$variable == Var.i, ]
#   YourFileName <- paste(Var.i, ".jpg", sep = "")
#   jpeg(file = YourFileName)
  print(ggplot(data = dfs.i, aes(x = ID.., y = value)) + geom_boxplot() + geom_point() + ggtitle(paste(Var.i, "_CTR", sep = "")) + theme(legend.position="none"))
  
}

```




#### Individual distribution by pers_var **EXPERIMENT**        
  
  
  
``` {r, fig.width=2.5, fig.height=2.5, message=FALSE, warning=FALSE}

AllVar4 <- unique(melt_persIDcs_e$variable)
for (i in 1:length(AllVar4))
{
  Var.i <- AllVar4[i]
  dfs.i <- melt_persIDcs_e[melt_persIDcs_e$variable == Var.i, ]
  print(ggplot(data = dfs.i, aes(x = ID.., y = value)) + geom_boxplot() + geom_point() + ggtitle(paste(Var.i, "_EXP", sep = "")) + theme(legend.position="none"))
  
}

```



------------


#### Per individual, all pers_vars **CONTROL**   


``` {r, fig.width=3, fig.height=3, message=FALSE, warning=FALSE}

ssize_c <- table(melt_persIDcs_c$ID..)/12 # sample size for each ind
ssize_c <-ssize_c [! ssize_c %in% 0] # to get rid of inds with n = 0

melt_persIDcs_co <- melt_persIDcs_c[order(melt_persIDcs_c$ID..),]


AllVar5 <- unique(melt_persIDcs_co$ID..)
for (i in 1:length(AllVar5))
{
  Id.i <-  ssize_c[i]
  Var.i <- AllVar5[i]
  dfs.i <- melt_persIDcs_co[melt_persIDcs_co$ID.. == Var.i, ]
  print(ggplot(data = dfs.i, aes(x = variable, y = value)) + geom_point() + geom_boxplot() + coord_flip() + ggtitle(paste(Var.i, "_CTR", "_N=", ssize_c [i], sep = "")) + theme(legend.position="none"))
  
}

```



#### Per individual, all pers_vars **EXPERIMENT**     


``` {r, fig.width=3, fig.height=3, message=FALSE, warning=FALSE}

ssize_e <- table(melt_persIDcs_e$ID..)/12 # sample size for each ind
ssize_e <-ssize_e [! ssize_e %in% 0] # to get rid of ind with n = 0

melt_persIDcs_eo <- melt_persIDcs_e[order(melt_persIDcs_e$ID..),]

AllVar6 <- unique(melt_persIDcs_eo$ID..)
for (i in 1:length(AllVar6))
{
  Id.i <-  ssize_e[i]
  Var.i <- AllVar6[i]
  dfs.i <- melt_persIDcs_eo[melt_persIDcs_eo$ID.. == Var.i, ]
  print(ggplot(data = dfs.i, aes(x = variable, y = value)) + geom_point() + geom_boxplot() + coord_flip() + ggtitle(paste(Var.i, "_EXP", "_N=", ssize_e [i], sep = "")) + theme(legend.position="none"))
  
}

```


#### Testing for interindividual differences **CONTROL** (only, unsuff n ind tested in exp)  


```{r, echo = TRUE, results="hide"}

# Inds with n = 1 excluded:
table(melt_persIDcs_c$ID..)/12 # sample size for each ind
melt_persIDcs_cs. <- melt_persIDcs_c[melt_persIDcs_c$ID.. != "241",]
melt_persIDcs_cs <- melt_persIDcs_cs.[melt_persIDcs_cs.$ID.. != "365",] # selected df with ind of n > 1

AllVard7 <- unique(melt_persIDcs_cs$variable)

for (i in 1:length(AllVard7))
{
  Var.i <- AllVard7[i]
  dfs.i <- melt_persIDcs_cs[melt_persIDcs_cs$variable == Var.i, ]
  res.i <- lm(value ~ ID.., data = dfs.i)
  print("######################################################################")
  print(Var.i)
  print(summary(res.i))
  print(anova(res.i))
}


```

```{r, echo = FALSE, message=FALSE, warning=FALSE}

library(knitr)

Var <- c("stan_nflo", "stan_nflochang", "stan_nouts", "stan_nins", "mean_durins", "tot_durins", "mean_durouts", "tot_durouts", "tot_durfor", "mov_spead", "mov_feroc", "stand_totdist")

Nind <-   c(1, 1, 2, 2, 0, 3, 2, 2, 4, 0, 0, 0)

Radj <- round(c(0.2316, 0.09081, 0.1859, 0.2754, 0.1101, 0.2557, 0.4368 ,0.3761, 0.3761, -0.06385,0.07824,-0.04119),2)
 
Pval <- round(c(0.0007999, 0.0677, 0.004046, 0.0001436, 0.04065, 0.0003175, 5.62e-08, 1.464e-06, 1.464e-06, 0.8923, 0.09258, 0.752), 3)
 
resdf1 <- as.data.frame(cbind(Nind, Radj, Pval))
resdf1$Var <- Var
resdf1 <-resdf1[,c(4,1:3)]
resdf1 <- resdf1[order(-resdf1$Nind),]

kable(resdf1, format = "html")

```
*Var - pers ariable; Nind - n individuals distinctly different (p < .05); Radj - R-adjsted of  the model; Pval - p-value of the model*    


####Relationships between the pers_var **CONTROL** (only ctr, unsuff n ind tested in exp)    


```{r, fig.width=12, fig.height=12}

panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(persIDcs_m[persIDcs_m$treat == "Ctr",3:14], lower.panel=panel.smooth, upper.panel = panel.cor, pch = 20)


```


##### Relationships between the pers_var **CONTROL, only vars with >2 n ind distinct in the ANOVA**


```{r, echo=FALSE, results="hide", fig.width=12, fig.height=12}

panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
 {
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(persIDcs_m[persIDcs_m$treat == "Ctr",c(4:7,12,13)], lower.panel=panel.smooth, upper.panel = panel.cor, pch = 20)

```


#### Clustering individuals - **CONTROL** - all pers_var - testing various distance and clustering methods


```{r, echo=TRUE, results="hide", fig.show="hide", message=FALSE, warning=FALSE}

library(pvclust)

persIDcs_mst <- t(persIDcs_m[persIDcs_m == "Ctr" , 3:14])
colnames(persIDcs_mst) <- unique(persIDcs_m[persIDcs_m == "Ctr" ,]$ID..)
persIDcs_mst_sc <- scale(persIDcs_mst)

methd <- c(rep("correlation", 5), rep("euclidean", 5), rep("manhattan", 5))
methcl <- c(rep(c("average", "ward.D", "ward.D2", "complete", "centroid"),3))

for (i in 1:length(methcl)) {
  methd.i <- methd [i]
  methcl.i <- methcl [i]
  fit.i <- pvclust(persIDcs_mst_sc, method.hclust=methcl[i],
               method.dist=methd[i], nboot = 1000, iseed = 20)
  print(plot(fit.i, main = "", hang = -1))
  print(pvrect(fit.i, alpha=.95))
}






```

#### Cophenetic correlations - to find the best method of dist and clust        

Distance matrices tested (order):: "correlation", "Euclidean", "Manhattan"          

```{r, echo=TRUE, message=FALSE, warning=FALSE}

library(proxy)

persIDcs_ms <- persIDcs_m[persIDcs_m == "Ctr" , 3:14]
colnames(persIDcs_ms) <- unique(persIDcs_m[persIDcs_m == "Ctr" ,]$ID..)
persIDcs_ms_sc <- scale(persIDcs_ms)

dist_cor <- proxy::dist(persIDcs_ms_sc, method = "correlation")
dist_eu <- proxy::dist(persIDcs_ms_sc, method = "Euclidean")
dist_mh <- proxy::dist(persIDcs_ms_sc, method ="Manhattan")


distmax <- rep(list (dist_cor, dist_eu, dist_mh),5)
cl_meth <- c(rep("average",3), rep("ward.D", 3), rep("ward.D2", 3), rep("complete", 3), rep("centroid", 3))

for (i in 1:length(distmax)) {
  hcl.i <- hclust(distmax[[i]], method = cl_meth[i])
  cop.i <- cophenetic(hcl.i)
  cor.i <- cor(distmax[[i]], cop.i)
  print(paste(cl_meth[i], " = ", cor.i, sep = ""))
}

```

###### Comp of the two trees with the best cophenetic correlations     
->  dist: "Euclidean"; clust: "average" and "centroid"

```{r, echo=FALSE, results="hide", fig.width=5, fig.height=5, message=FALSE, warning=FALSE}

library(pvclust)

persIDcs_mst <- t(persIDcs_m[persIDcs_m == "Ctr" , 3:14]) # data transposed - important for pvclust
colnames(persIDcs_mst) <- unique(persIDcs_m[persIDcs_m == "Ctr" ,]$ID..)
persIDcs_mst_sc <- scale(persIDcs_mst)

clust.av <- pvclust(persIDcs_mst_sc, method.hclust="average",
               method.dist="euclidean", nboot = 1000, iseed = 20)

plot(clust.av, hang = -1, main = "", ylab = "")
pvrect(clust.av, alpha = .95)



clust.com <- pvclust(persIDcs_mst_sc, method.hclust="complete",
               method.dist="euclidean", nboot = 1000, iseed = 20)

plot(clust.com, hang = -1, main = "", ylab = "")
pvrect(clust.com, alpha = .95)

```


### Testing cohenetic for the two above
```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(proxy)
persIDcs_ms <- persIDcs_m[persIDcs_m == "Ctr" , 3:14]
colnames(persIDcs_ms) <- unique(persIDcs_m[persIDcs_m == "Ctr" ,]$ID..)
persIDcs_ms_sc <- scale(persIDcs_ms)

dist_eu <- proxy::dist(persIDcs_ms_sc, method = "Euclidean")

set.seed(23235)

hc.av <- hclust(dist_eu, "average")
hc.com <- hclust(dist_eu, "complete")

cor(cophenetic(hc.av), cophenetic(hc.com))


```


### PCA

``` {r, echo=TRUE, message=FALSE, warning=FALSE}

library(FactoMineR)
library(factoextra)

# CONTROL
# data and PCA
df.pca<-persIDcs[persIDcs$treat == "Ctr",4:15]
df.pcaID <-persIDcs[persIDcs$treat == "Ctr",1]
df.pcatreat <-persIDcs[persIDcs$treat == "Ctr",2]

res.pca <- PCA(df.pca, graph = F)

# Eigenvalues and scree plot
eigv <- res.pca$eig
fviz_screeplot(res.pca)

# Correlations of variables with pc1 and pc2
fviz_pca_var(res.pca, axes = c(1,2), col.var = "cos2") +
scale_color_gradient2(low = "white", mid = "grey", high = "black", midpoint = 0.5) + theme_minimal()


# Contributions of particular variables into two first PCs 
fviz_pca_contrib(res.pca, choise = "var", axes = 1:2)


# EXPERIMENT
# data and PCA
df.pca.e<-persIDcs[persIDcs$treat == "Exp",4:15]
df.pcaID.e <-persIDcs[persIDcs$treat == "Exp",1]
df.pcatreat.e <-persIDcs[persIDcs$treat == "Exp",2]

res.pca.e <- PCA(df.pca.e, graph = F)

# Eigenvalues and scree plot
eigv.e <- res.pca.e$eig
fviz_screeplot(res.pca.e)

# Correlations of variables with pc1 and pc2
fviz_pca_var(res.pca.e, axes = c(1,2), col.var = "cos2") +
scale_color_gradient2(low = "white", mid = "grey", high = "black", midpoint = 0.5) + theme_minimal()


# Contributions of particular variables into two first PCs 
fviz_pca_contrib(res.pca.e, choise = "var", axes = 1:2)


```


``` {r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}

# Getting the PC scores for the individual observations - CONTROL
df.pca$pc1<-res.pca$ind$coord[,1] # to get the pc1 and pc2 scores!!!!
df.pca$pc2<-res.pca$ind$coord[,2] # to get the pc1 and pc2 scores!!!!
df.pca$pc3<-res.pca$ind$coord[,3] # to get the pc1 and pc2 scores!!!!
df.pca$pc4<-res.pca$ind$coord[,4] # to get the pc1 and pc2 scores!!!!
df.pca$pc5<-res.pca$ind$coord[,5] # to get the pc1 and pc2 scores!!!!
df.pca$ID..<- df.pcaID
df.pca$treat<- df.pcatreat


# Getting the PC scores for the individual observations - EXP
df.pca.e$pc1<-res.pca.e$ind$coord[,1] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc2<-res.pca.e$ind$coord[,2] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc3<-res.pca.e$ind$coord[,3] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc4<-res.pca.e$ind$coord[,4] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc5<-res.pca.e$ind$coord[,5] # to get the pc1 and pc2 scores!!!!
df.pca.e$ID.. <-df.pcaID.e
df.pca.e$treat <- df.pcatreat.e

```


#### PCA on selected variables

``` {r, echo=TRUE, message=FALSE, warning=FALSE}

library(FactoMineR)
library(factoextra)

# CONTROL
# data and PCA
names(persIDcs)
df.pca<-persIDcs[persIDcs$treat == "Ctr",6:15]
df.pcaID <-persIDcs[persIDcs$treat == "Ctr",1]
df.pcatreat <-persIDcs[persIDcs$treat == "Ctr",2]

res.pca <- PCA(df.pca, graph = F)

# Eigenvalues and scree plot
eigv <- res.pca$eig
fviz_screeplot(res.pca)

# Correlations of variables with pc1 and pc2
fviz_pca_var(res.pca, axes = c(1,2), col.var = "cos2") +
scale_color_gradient2(low = "white", mid = "grey", high = "black", midpoint = 0.5) + theme_minimal()


# Contributions of particular variables into two first PCs 
fviz_pca_contrib(res.pca, choise = "var", axes = 1:2)


# EXPERIMENT
# data and PCA
df.pca.e<-persIDcs[persIDcs$treat == "Exp",6:15]
df.pcaID.e <-persIDcs[persIDcs$treat == "Exp",1]
df.pcatreat.e <-persIDcs[persIDcs$treat == "Exp",2]

res.pca.e <- PCA(df.pca.e, graph = F)

# Eigenvalues and scree plot
eigv.e <- res.pca.e$eig
fviz_screeplot(res.pca.e)

# Correlations of variables with pc1 and pc2
fviz_pca_var(res.pca.e, axes = c(1,2), col.var = "cos2") +
scale_color_gradient2(low = "white", mid = "grey", high = "black", midpoint = 0.5) + theme_minimal()


# Contributions of particular variables into two first PCs 
fviz_pca_contrib(res.pca.e, choise = "var", axes = 1:2)


```


``` {r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}

# Getting the PC scores for the individual observations - CONTROL
df.pca$pc1<-res.pca$ind$coord[,1] # to get the pc1 and pc2 scores!!!!
df.pca$pc2<-res.pca$ind$coord[,2] # to get the pc1 and pc2 scores!!!!
df.pca$pc3<-res.pca$ind$coord[,3] # to get the pc1 and pc2 scores!!!!
df.pca$pc4<-res.pca$ind$coord[,4] # to get the pc1 and pc2 scores!!!!
df.pca$pc5<-res.pca$ind$coord[,5] # to get the pc1 and pc2 scores!!!!
df.pca$ID..<- df.pcaID
df.pca$treat<- df.pcatreat


# Getting the PC scores for the individual observations - EXP
df.pca.e$pc1<-res.pca.e$ind$coord[,1] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc2<-res.pca.e$ind$coord[,2] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc3<-res.pca.e$ind$coord[,3] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc4<-res.pca.e$ind$coord[,4] # to get the pc1 and pc2 scores!!!!
df.pca.e$pc5<-res.pca.e$ind$coord[,5] # to get the pc1 and pc2 scores!!!!
df.pca.e$ID.. <-df.pcaID.e
df.pca.e$treat <- df.pcatreat.e

```






#### Reapeted ANOVA on PCs
```{r,echo=FALSE, results="hide", message=FALSE, warning=FALSE }
df.pca.both <- rbind(df.pca.e, df.pca) # df needed for rep ANOVA
table(df.pca.both$treat, df.pca.both$ID..) # to check the matrix for the analysis

df.pca.both_rep.. <-df.pca.both[df.pca.both$ID.. != "143", ]
df.pca.both_rep. <-df.pca.both_rep..[df.pca.both_rep..$ID.. != "365", ]
df.pca.both_rep <- droplevels(df.pca.both_rep.)

table(df.pca.both_rep$treat, df.pca.both_rep$ID..) # to check the matrix for the analysis
names(df.pca.both)
library(plyr)

df.rep <- ddply(df.pca.both_rep, c("treat", "ID.."), summarise,
               pc1    = mean(pc1),
               pc2    = mean(pc2),
               pc3    = mean(pc3),
               pc4    = mean(pc4),
               pc5    = mean(pc5),
               # stan_nflo     = mean(stan_nflo),
               stan_nouts    = mean(stan_nouts),
               stan_nins     = mean(stan_nins),
               tot_durouts   = mean(tot_durouts),
               tot_durins    = mean(tot_durins),
               tot_durfor    = mean(tot_durfor),
               mov_spead     = mean(mov_spead)
)
               
               
# PC1           
aov.out = aov(pc1 ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
aov.out = aov(pc1 ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
#Plot
plotpc1<- ggplot(data = df.rep, aes(x = ID.., y = pc1, colour = treat)) + geom_point(colour = c(rep("gray33", 10), rep("steelblue2", 10)), size = c(rep(2, 10), rep(3, 10))) + ggtitle("PC1_undisturbed(gray)/experimental(blue)")    

# PC2
aov.out = aov(pc2 ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
aov.out = aov(pc2 ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
#Plot
plotpc2 <- ggplot(data = df.rep, aes(x = ID.., y = pc2, colour = treat)) + geom_point(colour = c(rep("gray33", 10), rep("steelblue2", 10)), size = c(rep(2, 10), rep(3, 10)))  + ggtitle("PC2_undisturbed(gray)/experimental(blue)")

grid.arrange(plotpc1, plotpc2, ncol = 2)


# Other variables

# stan_nflo
# aov.out = aov(stan_nflo ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = stan_nflo, colour = treat)) + geom_point() + ggtitle("means for tan_nflo")  
# 
# 
# # stan_nouts
# aov.out = aov(stan_nouts ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = stan_nouts, colour = treat)) + geom_point() + ggtitle("means for stan_nouts")  
# 
# # stan_nins
# aov.out = aov(stan_nins ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# aov.out = aov(stan_nins ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = stan_nins, colour = treat)) + geom_point() +
#           ggtitle("means for PC2")  
# 
# 
# # tot_durouts
# aov.out = aov(tot_durouts ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# aov.out = aov(tot_durouts ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = tot_durouts, colour = treat)) + geom_point() +
#           ggtitle("means for PC2")  
# 
# 
# # tot_durins
# aov.out = aov(tot_durins ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# aov.out = aov(tot_durins ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = tot_durins, colour = treat)) + geom_point() +
#           ggtitle("means for PC2")  
# 
# 
# # tot_durfor
# aov.out = aov(tot_durfor ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# aov.out = aov(tot_durfor ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = tot_durfor, colour = treat)) + geom_point() +
#           ggtitle("means for PC2")  
# 
# 
# # mov_spead
# aov.out = aov(mov_spead ~ treat + Error(ID../treat), data=df.rep); summary(aov.out)
# aov.out = aov(mov_spead ~ ID.. + Error(treat/ID..), data=df.rep); summary(aov.out)
# #Plot
# ggplot(data = df.rep, aes(x = ID.., y = mov_spead, colour = treat)) + geom_point() +
#           ggtitle("means for PC2")  
# 
# 

```

Results: all tests, p-values > .05    


#### Testing interindvidual differences with PC1 and PC2 **CONTROL only**
(presented output for pc1 only)

``` {r, echo=FALSE, message=FALSE, warning=FALSE}


# df.pca.both_rep # data frame
df.pcdf_c <- df.pca.both_rep[df.pca.both_rep$treat == "Ctr",]
df.pcdf_e <- df.pca.both_rep[df.pca.both_rep$treat == "Exp",]

give.n <- function(x){
   return(c(y = median(x)-.1, label = length(x)))
}

ggplot(data = df.pcdf_c, aes(x = ID.., y = pc1)) + geom_point() + geom_boxplot() + theme(legend.position="none") +
stat_summary(fun.data = give.n, geom = "text") 
 

ggplot(data = df.pcdf_e, aes(x = ID.., y = pc1)) + geom_point() + geom_boxplot() + theme(legend.position="none") +
stat_summary(fun.data = give.n, geom = "text") 

3

res.pc1 <- lm(pc1 ~ ID.., data = df.pcdf_c)
summary(res.pc1)
anova(res.pc1)
# 
# res.pc2 <- lm(pc2 ~ ID.., data = df.pcdf_c)
# summary(res.pc2)
# anova(res.pc2)

# res.pc1e <- lm(pc1 ~ ID.., data = df.pcdf_e)
# summary(res.pc1e)
# anova(res.pc1e)
# 
# res.pc2e <- lm(pc2 ~ ID.., data = df.pcdf_e)
# summary(res.pc2e)
# anova(res.pc2e)

```
